### 2021-01-08 ICLS_2020_Template.tex
Initial Overleaf Import

%-----------------------------------------------%
% Packages arranged by : Tsz Timmy Chan	        %
%                 Date : Apr 17th, 2020	        %
%	See ICLSpreamble.sty for details	        %
%	Uses biblatex-apa package for references    %
%-----------------------------------------------%
% File hosted on Overleaf and synced with Github

\documentclass{ICLSarticle} % Custom formatting file, based on standard article class
\usepackage{ICLSpreamble}   % Cusom commands defined here, as well as useful packages.

\ICLStitle{LaTeX, Overleaf, and RMarkdown Templates for ISLS Conferences: Increasing Collaboration, Reproducibility and Accessibility}
% Work Title Here.

% YOUR NAME(S) HERE 
\author{Author Name, Institution, Email \and Author Name 2, Institution, Email \and Author Name 3, Institution, Email%

\vspace{-9.5ex}} %removes spacing after authors

\addbibresource{ISLS2021.bib}		% Reference file, NO SPACES in the name

% CUSTOM PACKAGES GO HERE %
%\usepackage{LSRIcommon}
% CUSTOM PACKAGES GO HERE %
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\begin{document}

%====ABSTRACT AND KEYWORDS====%
\maketitle
% \begin{ICLSabstract}
% 	\input{abstract} % This template includes standalone package so this works nicely
% \end{ICLSabstract}
%	\ICLSkeywords{maximum five keywords should be included}

%====MAIN TEXT BELOW====%
\input{paper_contents} % inputs paper_content.tex

\printbibliography[title={References}]

\section{Acknowledgments}

Removed for review.

\end{document}