### 2021-01-08 abstract.tex

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{     }
\author{     }
\date{     }

\begin{document}
Learning sciences research has demonstrated the importance of social interactions during learning to help promote deep and meaningful understanding through a process of co-constructing knowledge, but homework reading is typically done as an isolated exercise. We have developed a social robot to provide social interactions during reading activities with middle-school children, and in this study report on how interacting with the robot affected the learning experience. Our thematic analysis describes both direct and indirect benefits from reading with the robot. We conclude with theoretical and practical implications of these results.

\end{document}